Nodal Discontinuous Galerkin Method for the Euler Equations in GR

### 2018-11-29 Documents/Euler/SamsTexFiles/Time-Step_Restriction_High-Order.tex
Updated dgEulerGR.tex document

We start by numerically computing the cell-average using quadrature with the Gauss-Lobatto quadrature rule. We assume that the DG approximation polynomial for the conserved variables is of order $k$, and that the order of the approximate solution is of order $k+d$, where $d$ is an integer that depends on the metric determinant. For the case of Cartesian coordinates, $d=0$ (because $\sqrtgm\sim x^{0}$), and for spherical-polar coordinates, $d=2$ (because $\sqrtgm\sim r^{2}$). Gauss-Lobatto integration will give an exact result if we choose a sufficiently high number, $M$, of quadrature points:
\begin{equation}
    2\,M-3\geq k+d\implies M\geq\f{k+d+3}{2}.
\end{equation}

@@

All of the terms in the square brackets are similar the the $\bH$ quantities in \citet{Qin2016}. We see that the expressions in the curly brackets are convex combinations (given a restriction on $\Delta t_{j}$), because the coefficients sum to unity. Since the quadrature weights are symmetric (so $w_{1}=w_{M}\equiv w_{GL}$), we find that the condition for $\ol{\bH}_{1}\in\cG$ is a time-step restriction:

@@

\Delta t_{j}<\ve\,\Delta x_{j}\,w_{GL}\,\text{min}\left(\f{1}{\alpha_{L}},\f{1}{\alpha_{H}}\right)

@@

Since we want a time-step that is constant for all elements, we choose:

@@

\Delta t<\ve\,w_{GL}\,\text{min}_{j}\left(\f{\Delta x_{j}}{\text{max}\left(\alpha_{L},\alpha_{H}\right)}\right).

@@

@@

\blue{NOTE:} We see the effect of the high-order approximation in the presence of the quadrature end-point weight $w_{GL}$. As the order increases, $w$ decreases, thus making a tighter restriction on the time-step.

>>>>

We start by numerically computing the cell-average using quadrature with the Gauss-Lobatto quadrature rule. We assume that the DG approximation polynomial for the conserved variables is of order $k$, and that the order of the approximate solution is of order $k+d$, where $d$ is an integer that depends on the metric determinant. For the case of Cartesian coordinates, $d=0$ (because $\sqrtgm\sim x^{0}$), and for spherical-polar coordinates in spherical symmetry, $d=2$ (because $\sqrtgm\sim r^{2}$). Gauss-Lobatto integration will give an exact result if we choose a sufficiently high number, $M$, of quadrature points:
\begin{equation}
    2\,M-3\geq k+d\implies M\geq\f{k+d+3}{2}.
\end{equation}

@@

All of the terms in the square brackets are similar the the $\bH$ quantities in \citet{Qin2016} and therefore belong to the set of admissible states, provided that:

@@

    \alpha_{L/H}=\alpha^{*}\geq\f{\left|v^{1}\right|\left(h+1-2\,h\,\tau\right)\,W^{2}+\sqrt{\tau^{4}\left(h-1\right)^{2}+\tau^{2}\left(h-1\right)\left(h+1-2\,h\,\tau\right)}}{W^{2}\left(h+1-2\,h\,\tau\right)+\tau^{2}\left(h-1\right)},

@@

where $h$ is the relativistic specific enthalpy and

@@

    \tau\equiv\f{\Gamma-1}{\Gamma},

@@

where $\Gamma$ is the adiabatic index.

We see that the expressions in the curly brackets are convex combinations (given a restriction on $\Delta t_{j}$), because the coefficients sum to unity. Since the quadrature weights are symmetric (so $w_{1}=w_{M}\equiv w_{GL}$), we find that the condition for $\ol{\bH}_{1}\in\cG$ is a time-step restriction:
\begin{equation}
\Delta t_{j}<\ve\,\Delta x_{j}\,w_{GL}\,\text{min}\left(\f{1}{\alpha_{L}},\f{1}{\alpha_{H}},\f{1}{\alpha^{*}}\right)
\end{equation}
Since we want a time-step that is constant for all elements, we choose:
\begin{equation}
\Delta t<\ve\,w_{GL}\,\text{min}_{j}\left(\f{\Delta x_{j}}{\text{max}\left(\alpha_{L},\alpha_{H},\alpha^{*}\right)}\right).
\end{equation}

@@

\blue{NOTE:} We see the effect of the high-order approximation in the presence of the quadrature end-point weight $w_{GL}$. As the order increases, $w_{GL}$ decreases, thus making a tighter restriction on the time-step.
